Simulations of Ballot Polling Risk-Limiting Audits

### 2023-02-19 providence/sections/intro.tex
re: beta

As a formalized approach to the examination of the evidence supporting vote tabulation, an RLA is an important part of an evidence-based election. When correctly used, it serves to greatly improve the trustworthiness of an election.

@@

\subsubsection{The Workflow of a Ballot Polling RLA}
\label{sec:polling}
%Ballot and batch comparison RLAs compare information gathered through manual examination of the sampled ballots with the corresponding information claimed by the voting system. For example, the information could be the vote on a particular ballot (stored by the voting system as a CVR or Cast Vote Record) or the tally of a particular box of ballots. 
A ballot polling RLA \cite{RLA} is based on the manual examination of the sampled ballots, and does not require any information from the tabulating system other than the tally. 
%, on the other hand, simply tallies the sampled ballots. It requires a much larger number of ballots than a ballot comparison RLA, but tends to be more feasible because it does not require any additional functionality of the voting system; in particular it does not require the system to generate or provide CVRs. 
More detail about the storage of ballots is required, however: a complete ballot manifest (a list of ballot storage containers and the number of ballots in each) which enables the creation of a well defined list of the ballots and their locations (the fifth ballot in box number 20, for example) to enable the sampling of specific ballots from the list.

@@

\subsubsection{The Audit Model}
An audit is typically defined as a binary hypothesis test. If the null, $H_0$, is defined as the incorrect outcome hardest to detect (generally a tie, see \cite{Bayesian-RLA}, we have the following definition of an RLA.

@@

We will refer to $Pr[\mathcal{A}(X) = \text{Correct} \,|\, H_0]$ as the {\em risk} of the audit (though, technically, it is the maximum risk, corresponding to the election outcome hardest to detect). Every time an audit stops, it incurs a contribution to its risk. Audit design determines how the risk contributions are distributed across rounds, so long as the overall risk is limited by $\alpha$.

@@

{\bf Full Hand Count:} It is worth noting here that if election officials do not have a plan for when they will move to a full hand count, at least a $1-\alpha$ fraction of the audits of an incorrect outcome will {\em never} stop, and election officials will keep drawing new rounds of ballots. See Figure \ref{fig:prov-risk} in section \ref{sec:sims} for the results of 10,000 simulations of \Providence on tied elections (risk limit of 0.1, across election margins). Observe that, for each margin studied, more than 90\% of the audits do not stop after 5 rounds.

@@

To avoid a scenario where election officials fruitlessly draw round after round hoping the audit will stop when the election outcome is incorrect, it might be worthwhile for election officials to determine, before the audit begins, the latest date by which either the audit stops or a hand count must begin to be completed before certification.

{\bf The Adversary in an RLA:} The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, make the audit declare an incorrect outcome as correct more often than implied by the risk limit. An audit is an RLA only if there is a proof of its risk-limiting property. Thus, the adversary would need to, at the very least, invalidate an assumption of the proof to obtain:  
$$
Pr[\mathcal{A}(X) 
= \text{Correct} \,|\, H_0] > \alpha
$$

>>>>

As a formalized approach to the examination of evidence supporting vote tabulation, an RLA is an important part of an evidence-based election. When correctly implemented, it serves to vastly improve the trustworthiness of the election.

@@

\subsubsection{The Workflow of a Ballot Polling RLA}
\label{sec:polling}
%Ballot and batch comparison RLAs compare information gathered through manual examination of the sampled ballots with the corresponding information claimed by the voting system. For example, the information could be the vote on a particular ballot (stored by the voting system as a CVR or Cast Vote Record) or the tally of a particular box of ballots. 
A ballot polling RLA \cite{RLA} is based on manual examination of the sampled ballots, and does not require any information from the tabulating system other than the tally. 
%, on the other hand, simply tallies the sampled ballots. It requires a much larger number of ballots than a ballot comparison RLA, but tends to be more feasible because it does not require any additional functionality of the voting system; in particular it does not require the system to generate or provide CVRs. 
More detail about the storage of ballots is required, however: a complete ballot manifest (a list of ballot storage containers and the number of ballots in each) which enables the creation of a well defined list of the ballots and their locations (the fifth ballot in box number 20, for example) to enable the sampling of specific ballots from the list.

@@

\subsubsection{The Audit Model}
An audit is typically defined as a binary hypothesis test. If the null, $H_0$, is defined as the incorrect outcome hardest to detect (generally a tie, see \cite{Bayesian-RLA}), we have the following definition of an RLA.

@@

We will refer to $Pr[\mathcal{A}(X) = \text{Correct} \,|\, H_0]$ as the {\em risk} of the audit\footnote{Technically, it is the maximum risk, corresponding to the election outcome hardest to detect.}. Every time an audit stops, it incurs a contribution to its risk. Audit design determines how the risk contributions are distributed across rounds, so long as the overall risk is limited by $\alpha$.

@@

{\bf Full Hand Count:} It is worth noting here that if election officials do not have a plan for when they will move to a full hand count, and the election outcome is incorrect, at least a fraction $(1-\alpha)$ of the audits will {\em never} stop, and they will keep drawing new rounds of ballots. See Figure \ref{fig:prov-risk} in section \ref{sec:sims} for the results of 10,000 simulations of \Providence on tied elections (risk limit of 0.1, across election margins). Observe that, for each margin studied, more than 90\% of the audits do not stop after 5 rounds.

@@

To avoid a scenario where election officials fruitlessly draw round after round hoping the audit will stop when the election outcome is incorrect, it might be worthwhile for election officials to determine, before the audit begins, the latest date by which either: the audit stops, or a hand count begins, so as to be completed before certification.

{\bf The Adversary in an RLA:} The goal of the adversary is to increase the true risk beyond the declared risk limit---that is, make the audit declare an incorrect outcome as correct with a chance larger than the risk limit. An audit is an RLA only if there is a proof of its risk-limiting property. Hence, at the very least, the adversary would need to invalidate an assumption of the proof to obtain:  
$$
Pr[\mathcal{A}(X) 
= \text{Correct} \,|\, H_0] > \alpha
$$

### 2023-02-19 providence/sections/model.tex
re: beta

An example of the above use is \cite{RLA}. In the more recent literature, for example \cite{bravo}, ballot polling audits do not include this possibility (i.e. they set $\beta=0$) in order to give maximum flexibility to election officials in choosing when to proceed to a full manual count. 
%In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count. This option is generally not incorporated \cite{bravo}. 
% Election officials would typically decide to perform a full manual hand count if the audit does not stop in spite of drawing a large number of ballots, typically over multiple rounds.

A manual hand count presents significant logistical challenges. Left to election officials, the decision would be influenced by the certification deadline, the estimated number of human hours required for another round, the logistical costs of a full hand count, and the impact of any decision on citizen confidence. A better workload model would better inform the decision. %The only disadvantage of leaving the decision to election officials would be when the outcome were incorrect, and election officials would continue drawing rounds hoping that they would not need to escalate to a full hand count. 
Audits going towards statutory or legal requirements would need to be completed by the certification deadline\footnote{Pilot audits, performed after certification, usually end after a single round or fixed number of rounds, providing a measured risk (the statistical p-value) at the end of the round, and no decision regarding a full hand count needs to be made.}; this provides an incentive to go to a hand count. It is worth considering bring back a small value of $\beta$ into the stopping condition, or requiring election officials to declare beforehand when they would stop drawing ballots and go to a full hand count.

>>>>

An example of the above use is \cite{RLA}. In the more recent literature, for example \cite{bravo}, ballot polling audits do not include this possibility (i.e. they set $\beta=0$) in order to give maximum flexibility to election officials in choosing when to proceed to a full manual count. For a given election, $\sigma$ is a function only of $k$, and increases with $k$. Any non-zero value of $\beta$ would reduce the value $\sigma$ is being compared with, and allow $\mathcal{A}$ to declare \emph{Correct} for lower values of $k$. Thus $\beta = 0$ is the most stringent of this form of audit. The choice of going to a full hand count when $\beta=0$ will only reduce the risk because it never results in an error (by definition, the true election outcome is one that would be obtained by a full manual hand count of the ballots). 
%In principle, the stopping condition could indicate a third option too: (c) the audit proceeds to a full manual hand count. This option is generally not incorporated \cite{bravo}. 
% Election officials would typically decide to perform a full manual hand count if the audit does not stop in spite of drawing a large number of ballots, typically over multiple rounds.

A manual hand count presents significant logistical challenges. The decision to move to one would be influenced by the certification deadline\footnote{As mentioned earlier, audits going towards statutory or legal requirements would need to be completed by the certification deadline. Pilot audits, performed after certification, usually end after a single round or fixed number of rounds, providing a measured risk (the statistical p-value) at the end of the round, and no decision regarding a full hand count needs to be made.}, the estimated number of human hours required for another round, the logistical costs of a full hand count, and the impact of any decision on citizen confidence. A better workload model would better inform the decision. %The only disadvantage of leaving the decision to election officials would be when the outcome were incorrect, and election officials would continue drawing rounds hoping that they would not need to escalate to a full hand count. 
Election officials should announce ahead of time their plans for when they would abort the audit procedure and go to a full hand count, justifying it with an eye towards completing certification requirements. This would provide a more transparent process and protect them from political pressures. See section \ref{sec:sims} for a discussion of the considerations informing the choice between drawing more rounds and performing a full hand count for the \Providence audit and large round size.
